Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-26 paper.tex
Minor updates

  workflow \emph{engine}. \note{Marc-e}{automation?}

@@

Regarding extensibility, most architectures are overall comparable,
except nested workflows (with service invocation) which is
significantly behind. This is explained by the complexity of the
nested workflow architecture, with 2 infrastructures and 2 workflow
services. Sub-tasking and workflow import (with service invocation)
perform a bit better when integrating new engines (\texttt{E$_1$}),
which is interesting given the current profusion of engines with
different characteristics.  However, sub-tasking, workflow import
(with service invocation) and tight integration perform worse than the
others when adding engine versions (\texttt{E$_2$}) due to the need to
update infrastructure \note{Marc-e}{confusion with
  physical?}(sub-tasking), science gateway (tight integration), or
workflow converter (workflow import). Workflow import performs worse
than the others when integrating new workflows (\texttt{E$_3$})
because of the language conversion step. All architectures except
nested workflows perform the same when integrating new infrastructures
(\texttt{E$_4$}).

The pool architecture is overall the most scalable, which is not a
surprise since it was designed precisely for scalability. Tight
integration is the least scalable and all the other architectures
perform the same. The overall scalability score, however, should not
mask the unique characteristics of architectures regarding this
criterion. Nested workflows are the only architecture that can easily
accommodate distributed workflow execution, which can be critical in
some cases. At the same time, the scheduling constraints created by
the sub-tasking and nested workflow architectures could well become
showstoppers depending on the type of the targeted
infrastructure. Non-reliable infrastructures, for example, could
hardly cope with workflow engines being wrapped in computing tasks as
done in sub-tasking \note{Marc-e}{free for all? WE dev involved?}. The availability of multiple engine instances
could also become a critical feature for science gateways with
important workloads, which would favor sub-tasking and pool.

@@

>>>>

  workflow \emph{engine}. 

@@

Regarding extensibility, most architectures are overall comparable,
except nested workflows (with service invocation) which is
significantly behind. This is explained by the complexity of the
nested workflow architecture, with 2 infrastructures and 2 workflow
services. Sub-tasking and workflow import (with service invocation)
perform a bit better when integrating new engines (\texttt{E$_1$}),
which is interesting given the current profusion of engines with
different characteristics.  However, sub-tasking, workflow import
(with service invocation) and tight integration perform worse than the
others when adding engine versions (\texttt{E$_2$}) due to the need to
update infrastructure (sub-tasking), science gateway (tight
integration), or workflow converter (workflow import). Workflow import
performs worse than the others when integrating new workflows
(\texttt{E$_3$}) because of the language conversion step. All
architectures except nested workflows perform the same when
integrating new infrastructures (\texttt{E$_4$}).

The pool architecture is overall the most scalable, which is not a
surprise since it was designed precisely for scalability. Tight
integration is the least scalable and all the other architectures
perform the same. The overall scalability score, however, should not
mask the unique characteristics of architectures regarding this
criterion. Nested workflows are the only architecture that can easily
accommodate distributed workflow execution, which can be critical in
some cases. At the same time, the scheduling constraints created by
the sub-tasking and nested workflow architectures could well become
showstoppers depending on the type of the targeted
infrastructure. Non-reliable infrastructures, for example, could
hardly cope with workflow engines being wrapped in computing tasks as
done in sub-tasking. The availability of multiple engine instances
could also become a critical feature for science gateways with
important workloads, which would favor sub-tasking and pool.

@@

% Literature on SG architecture Very few works go beyond the
% description of architectures of particular systems. The work in
% http://link.springer.com/chapter/10.1007/11596141_3 discusses
% service-oriented architectures for science gateways, with a workflow
% engine. The work in
% http://docs.lib.purdue.edu/cgi/viewcontent.cgi?article=1414&context=nanopub
% highlights 4 TeraGrid science gateways (GridChem, LEAD, nanoHUB,
% CABIG). Architectures are described but there is no real focus on
% workflow engines. GridChem doesn't include any workflow engine,
% which btw make it relevant for integration via sub-tasking. LEAD has
% a workflow engine.
